A White Noise Approach to Evolutionary Ecology

### 2020-04-28 Current Version/Current-Version.tex
reorganize

\begin{equation}
N(0)=\int_{-\infty}^{+\infty}\nu(x,0)dx<+\infty,
\end{equation} \begin{equation}
-\infty<\bar x(0)=\int_{-\infty}^{+\infty}xp(x,0)dx<+\infty,
\end{equation} \begin{equation}
\sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
\end{equation}

where \(\bar x(t)\) and \(\sigma^2(t)\) are respectively the mean trait
and phenotypic variance at time \(t\geq0\). Second, we assume selection
is determined by the growth rate \(m(h,x)\) that is differentiable with
respect to both arguments and is bounded above by some
\(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We also
assume \((1+h)C\leq m(h,x)\) for some constant \(C>0\) so that
\(m(h,x)\) is linear or more than linear in \(h\). We denote by \(K\) an
operator that takes functions such as \(\nu(x,t)\) as its argument and
returns other functions that may still depend on the spatial variable
\(x\) and the temporal variable \(t\). For our application we will be
focused on operators \(K\) that can be written as
\((K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x-y)\nu(y,t)dy\) for some
non-negative and bounded function \(\kappa\). We set \(h=(K\nu)(x,t)\)
to account for nonlocal effects of the abundance density \(\nu(x,t)\) on
the fitness of individuals with trait value \(x\) (Volpert, 2014). As an
example of a nonlocal effect, consider a population experiencing
intraspecific competition for a common resource such that the
competitive interactions are not mediated by any set of traits. In this
case all individuals in the population effect the fitness of one another
even though they may exhibit a diverse range of trait values. The term
nonlocal comes from interpreting the trait value as spatial location.
From here on we abbreviate \(m(h,x)=m\big((K\nu)(x,t),x\big)\) to just
\(m(\nu,x)\). Third, we assume mutation is captured by diffusion with
coefficient \(\frac{\mu}{2}\). With these technicalities addressed we
model the demographic dynamics of a population and the dynamics of a
quantitative character simultaneously by the PDE

\begin{equation}\label{eq1}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
\end{equation}

with the initial condition \(\nu(x,0)\) described above. Equation
(\ref{eq1}) is a semilinear parabolic PDE and is referred to as a scalar
reaction-diffusion equation (Evans, 2010; Zheng, 2004). To ensure
existence and uniqueness of solutions to (\ref{eq1}), we further assume
that for each \(M>0\) there exists a constant \(L_M>0\) depending on
\(M\) such that
\(\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M\) implies

@@

\begin{equation}\label{special_m}
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
\end{equation}

where \(r\in\mathbb{R}\) is the intrinsic rate of growth in the absence
of abiotic stabilizing selection, \(a>0\) is the strength of abiotic
stabilizing selection around the phenotypic optimum
\(\theta\in\mathbb{R}\), \(c>0\) captures the sensitivity of fitness to
intraspecific competition and \(\kappa(x-y)=1\) so that competitive
interactions cause the same reduction in fitness regardless of trait
value. This exemplary fitness function has a few convenient properties.
First, the effect of competition induces a local carrying capacity on
the population, leading to a finite equilibrium abundance over bounded
subsets of \(\mathbb{R}\). Second, abiotic selection prevents the
abundance density from diffusing too far from the abiotic optimum,
leading to a finite equilibrial abundance across all of \(\mathbb{R}\),
\(\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})\),
finite equilibrial mean trait \(\lim_{t\to\infty}\bar x(t)=\theta\) and
finite equilibrial phenotypic variance
\(\lim_{t\to\infty}\sigma^2(t)=\sqrt{\frac{\mu}{a}}\), so long as
\(\bar x(0)\in\mathbb{R}\), \(\sigma^2(0)\in[0,+\infty)\),
\(N(0)\in(0,+\infty)\) and \(r>\frac{1}{2}\sqrt{a\mu}>0\). We prove
these statements in SM \S\ref{equilib}. Note in particular the
coincidence of equilibrial phenotypic variance predicted by this model
and that predicted by the classic Gaussian allelic model of quantitative
genetics (Bürger, 2000; Johnson and Barton, 2005; Lande, 1975). In
\S\ref{coev} and SM \S\ref{diffuse} we find this fitness function is
closely related to a fitness function derived from niche theory that we
use to model coevolution driven by resource competition.

@@

\begin{equation}
\|f\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)},
\end{equation}

and make use of the convention \(f=g\) if \(\|f-g\|_t=0\) for all
\(t\geq0\). Later, when we investigate stochastic dynamics of the
abundance density \(\nu(x,t)\), we will find
\(\sqrt\nu\in\mathscr{N}_2\). This enables us to utilize the heuristics
developed in this section for the derivation of SDE describing the
stochastic dynamics of \(N(t)\), \(\bar x(t)\) and \(\sigma^2(t)\).

@@

\begin{equation}\label{exp_WN}
\mathbb{E}\big(\mathbf W_t(f)\big)=\mathbb{E}\big(\mathbf W_t(g)\big)=0,
\end{equation} \begin{equation}\label{cov_WN}
\mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\mathbb{E}\left(\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds\right),
\end{equation}

where \(t_1\wedge t_2=\min(t_1,t_2)\) and \(\mathbb{C}\) denotes
covariance with respect to the underlying probability space. In
particular, denoting \(\mathbb{V}\) the variance operator with respect
to the underlying probability space, we have
\(\mathbb{V}\big(\mathbf W_t(f)\big)=\|f\|_t^2\) for all \(t\geq0\) and
\(f\in\mathscr{N}_2\). The operators \(\mathbb{E}\) and \(\mathbb{C}\)
are to be distinguished from \(\bar f(t)\) and \(\mathrm{Cov}_t(f,g)\)
which denote expectation and covariance with respect to phenotypic
diversity at time \(t\geq0\).

@@

\begin{equation}
\|I_D\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_D^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
\end{equation}

where \(|D|\) denotes the length of \(D\), we have
\(I_D\in\mathscr{N}_2\). Thus, using equations (\ref{exp_WN}) and
(\ref{cov_WN}) and adopting the informal notation introduced in equation
(\ref{informal}), we can write the following

@@

\begin{equation}
\mathbb{E}\left(\int_0^td\hat{\mathbf W}_s(f)\right)=0, \ \mathbb{C}\left(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\right)=t_1\wedge t_2
\end{equation}

and in particular, as a function of \(t\),
\(\int_0^td\hat{\mathbf W}_s(f)\) is a standard Brownian motion for any
\(f\in\mathscr{N}_2\). Hence, \(d\hat{\mathbf W}_t(f)\) is analogous to
the traditional shorthand used to denote stochastic differentials. Thus,
equation (\ref{cov_WN_xi}) effectively extends Itô's multiplication
table to:

@@

\emph{Branching Brownian motion (BBM):} A BBM tracks individuals
navigating \(d\)-dimensional Euclidean space that reproduce and senesce
between exponentially distributed intervals. Unlike other stochastic
processes that take values in \(\mathbb{R}^d\), BBM takes values in the
set of \emph{non-negative finite measures} over \(\mathbb{R}^d\).
Intuitively, one can think of a finite measure as a function that maps
subsets of \(\mathbb{R}^d\) to real numbers. To be formal, we only
consider the Borel subsets of \(\mathbb{R}^d\) corresponding to the
Euclidean metric, but understanding this technicality is not crucial to
our discussion. In particular, denoting \(X_t\) a BBM, for a subset
\(D\subset\mathbb{R}^d\), \(X_t(D)\) returns the (random) number of
individuals alive within the region \(D\) at time \(t\geq0\). The BBM
has three main components:

@@

\emph{Interacting superprocesses:} In the above subsection we reviewed
the diffusion limit of an especially tractable measure-valued branching
process. However, we found the simplicity of this process restricts us
from modelling nonlocal effects on the fitness of individuals. Here, we
discuss superprocesses that account for such effects. The existence of
diffusion limits for a class of measure-valued branching processes
involving nonlocal effects has been treated by Méléard and Roelly (1993,
1992). The interactions can manifest as dependencies of the spatial
movement or reproductive law of individuals on their position \(x\) and
the state of the whole population described by \(X_t\). An important
result of Méléard and Roelly (1993, 1992) is a theorem that provides
sufficient conditions to construct a sequence of rescaled measure-valued
branching processes that converge to a generalization of the
Dawson-Watanabe superprocess that includes interactions. The rescaling
is analogous to that described above for non-interacting Dawson-Watanabe
superprocesses, but now the reproductive law described by
\(\mathscr{W}(X_t,x)\) and \(V(X_t,x)\), branching rate
\(\lambda(X_t,x)\) and diffusion parameter \(\sqrt{\mu(X_t,x)}\) are
allowed to depend on the whole population \(X_t\) and individual
location \(x\). In Figure \ref{rescaled} we demonstrate this rescaling
in discrete time for a population experiencing stabilizing selection and
logistic growth. Since time is discretized, the process we simulate is
formally a branching random walk. For further details on our simulation
see SM \S\ref{numerical}.

@@

\begin{multline}
N(t)=N(0)+\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
=N(0)+\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
\end{multline}

\begin{equation}
\bar m(t)=\frac{1}{N(t)}\int_\mathbb{R}m(\nu,x)\nu(x,t)dx,
\end{equation}

and

@@

\begin{subequations}\label{gen_eqns}
\begin{equation}\label{N}
dN(t)=\bar m(t)N(t)dt+\sqrt{V N(t)}dW_1(t),
\end{equation}
\begin{equation}\label{xbar_gen}
d\bar x(t)=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)dt+\sqrt{V\frac{\sigma^2(t)}{N(t)}}dW_2(t),
\end{equation}
\begin{equation}\label{sig2_gen}
d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
\end{equation}
\end{subequations}

@@

\begin{equation}\label{covxm}
\mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right),
\end{equation} \begin{equation}
\mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
\end{equation}

and \(\overline{(x-\bar x)^4}=3\sigma^4\). In particular, these results
imply

@@

@@

\begin{equation}
\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
\end{equation}

@@

To map the degree of niche overlap to fitness, we assume competition
between individuals with niche locations \(x_i\) and \(x_j\) decreases
the expected reproductive output for the individual in species \(i\) at
the rate \(c_i\mathcal{O}_{ij}(x_i-x_j)\) for some \(c_i>0\). We refer
to \(c_i\) as the sensitivity to competition for species \(i\). The term
\(c_i\mathcal{O}_{ij}(x_i-x_j)\) coincides with a special case of a term
used to capture competition in Dawson's geostochastic logistic model, an
SPDE model developed to study the combined effects of demographic
stochasticity, spatial dispersion and locally finite carrying capacity
(Dawson, 1978). In relation to the example fitness function discussed in
\S\ref{deterministic}, consider
\(\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)\). In SM \S\ref{diffuse} we
combine our treatment of resource competition with equations (\ref{N}),
(\ref{xbarfinal}) and (\ref{Gfinal}) to find

@@

@@

@@

\begin{multline}
\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy \\ =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
\end{multline}

@@

\emph{Metric of pairwise coevolution:} Below we investigate the
correspondence of interaction intensity and coevolutionary change.
However, we can already identify one major discrepancy; \(\alpha_{ij}\)
is maximized when \(\bar x_i=\bar x_j\), but \(\beta_{ij}=0\) under the
same condition. We therefore include in our metric of selection the
standardized stabilizing selection gradient \(\gamma\) which measures
the effect of stabilizing or disruptive selection on phenotypic variance
(Lande and Arnold, 1983). In SM \S\ref{diffuse} we show that the
standardized stabilizing selection gradient induced by species \(j\) on
species \(i\) can be computed as

@@

\emph{Covariance of selection and competition as a function of
diversity:} We now make use of the expressions derived for competition
coeffecients and selection gradients to investigate their relationship.
As a first pass, let us assume all model parameters are equivalent
across species and that each species has the same abundance and trait
variance. Let us further assume that richness \(S\) is large and the
distribution of mean trait values is normal with mean \(\bar{\bar x}\),
variance \(V_{\bar X}\) and density \(f_{\bar X}\). Such assumptions are
typical when deriving analytical results in the field of theoretical
coevolutionary community ecology (Nuismer et al., 2012, 2018). If
\(\bar{\bar x}\) is near \(\theta\) and \(V_{\bar X}\) is much smaller
than \(|2\frac{R}{a}-G-\eta|\), then we may approximate \(r_i\) with

@@

{\centering \includegraphics[width=1\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/cov}

@@

{\centering \includegraphics[width=0.6\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/.A White Noise Approach To Evolutionary Ecology/Cαℭp}

@@

To understand patterns found in ecological communities considerations
must push beyond microevolutionary and contemporary ecological processes
and consider the macroevolutionary dynamics of ancestral lineages
leading to extant populations. Using sub-alpine flower communities as an
example, one can observe a very strict ordering of phenology across a
broad geographic range. In particular, whether in the Colorado Rocky
mountains (such as Gothic, Colorado) or on an outlier of the Idaho
batholith (such as Kamiak butte near Palouse, Washington), one will
almost surely observe a very conspicuous order of flowers emerging in
early spring: at the very beginning of the season blooms \emph{Claytonia
lanceolata} followed by \emph{Erythronium grandiflorum} which precedes
\emph{Delphinium nuttallianum} (B. Week, personal observations). If
contemporary phenological coevolution is rampant, why should this
pattern be so well preserved across a thousand miles of rugged and
diverse terrain? Although it would be exciting to find that these
species repeatedly coevolved this pattern in each location, a more
parsimoniuous hypothesis suggests the phenology and physiology of these
species slowly evolved independently over macroevolutionary time scales
to take advantage of the specific conditions available within each of
these windows of the flowering season. However, this could not have
carried out in the Rocky mountains since this terrain only became
habitable just over ten thousand years ago as the glaciers of the
Pleistocene began to retreat (Paul CaraDonna, personal communications).
Hence, given these considerations, it appears that an understanding of
early season phenology patterns should focus on how these communities
are assembled as opposed to contemporary evolutionary dynamics. Indeed,
recent work testing models of phylogeography ignores the potential for
contemporary evolution and instead suggests alpine flower communities
tend to follow neutral assembly where flowers merely compete for who can
disperse to new habitat first, as opposed to a selective process where a
regional species pool is filtered for those species adapted to the newly
available habitat (Marx et al., 2017).

Of course microevolutionary and ecological dynamics are not completely
irrelevant for understanding patterns in communities that are primarily
structured by deep evolutionary processes. In particular,
macroevolutionary trait evolution is simply the aggregation of
microevolutionary change occuring over large spans of time. This
suggests a road forward to connect the theory we have introduced to
models of macroevolutionary trait evolution.

>>>>

\begin{equation}
N(0)=\int_{-\infty}^{+\infty}\nu(x,0)dx<+\infty,
\end{equation} \begin{equation}
-\infty<\bar x(0)=\int_{-\infty}^{+\infty}xp(x,0)dx<+\infty,
\end{equation} \begin{equation}
\sigma^2(0)=\int_{-\infty}^{+\infty}(x-\bar x(0))^2p(x,0)dx<+\infty,
\end{equation} where \(\bar x(t)\) and \(\sigma^2(t)\) are respectively
the mean trait and phenotypic variance at time \(t\geq0\). Second, we
assume selection is determined by the growth rate \(m(h,x)\) that is
differentiable with respect to both arguments and is bounded above by
some \(r\in\mathbb{R}\) for all \(x\in\mathbb{R}\) and \(h\geq0\). We
also assume \((1+h)C\leq m(h,x)\) for some constant \(C>0\) so that
\(m(h,x)\) is linear or more than linear in \(h\). We denote by \(K\) an
operator that takes functions such as \(\nu(x,t)\) as its argument and
returns other functions that may still depend on the spatial variable
\(x\) and the temporal variable \(t\). For our application we will be
focused on operators \(K\) that can be written as
\((K\nu)(x,t)=\int_{-\infty}^{+\infty}\kappa(x-y)\nu(y,t)dy\) for some
non-negative and bounded function \(\kappa\). We set \(h=(K\nu)(x,t)\)
to account for nonlocal effects of the abundance density \(\nu(x,t)\) on
the fitness of individuals with trait value \(x\) (Volpert, 2014). As an
example of a nonlocal effect, consider a population experiencing
intraspecific competition for a common resource such that the
competitive interactions are not mediated by any set of traits. In this
case all individuals in the population effect the fitness of one another
even though they may exhibit a diverse range of trait values. The term
nonlocal comes from interpreting the trait value as spatial location.
From here on we abbreviate \(m(h,x)=m\big((K\nu)(x,t),x\big)\) to just
\(m(\nu,x)\). Third, we assume mutation is captured by diffusion with
coefficient \(\frac{\mu}{2}\). With these technicalities addressed we
model the demographic dynamics of a population and the dynamics of a
quantitative character simultaneously by the PDE

\begin{equation}\label{eq1}
\frac{\partial}{\partial t}\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}\nu(x,t)
\end{equation} with the initial condition \(\nu(x,0)\) described above.
Equation (\ref{eq1}) is a semilinear parabolic PDE and is referred to as
a scalar reaction-diffusion equation (Evans, 2010; Zheng, 2004). To
ensure existence and uniqueness of solutions to (\ref{eq1}), we further
assume that for each \(M>0\) there exists a constant \(L_M>0\) depending
on \(M\) such that
\(\int_\mathbb{R}\nu_1(x)dx,\int_\mathbb{R}\nu_2(x)dx\leq M\) implies

@@

\begin{equation}\label{special_m}
m(\nu,x)=r-\frac{a}{2}(\theta-x)^2-c\int_{-\infty}^{+\infty}\nu(y,t)dy
\end{equation} where \(r\in\mathbb{R}\) is the intrinsic rate of growth
in the absence of abiotic stabilizing selection, \(a>0\) is the strength
of abiotic stabilizing selection around the phenotypic optimum
\(\theta\in\mathbb{R}\), \(c>0\) captures the sensitivity of fitness to
intraspecific competition and \(\kappa(x-y)=1\) so that competitive
interactions cause the same reduction in fitness regardless of trait
value. This exemplary fitness function has a few convenient properties.
First, the effect of competition induces a local carrying capacity on
the population, leading to a finite equilibrium abundance over bounded
subsets of \(\mathbb{R}\). Second, abiotic selection prevents the
abundance density from diffusing too far from the abiotic optimum,
leading to a finite equilibrial abundance across all of \(\mathbb{R}\),
\(\lim_{t\to\infty}N(t)=\tfrac{1}{c}(r-\tfrac{1}{2}\sqrt{a\mu})\),
finite equilibrial mean trait \(\lim_{t\to\infty}\bar x(t)=\theta\) and
finite equilibrial phenotypic variance
\(\lim_{t\to\infty}\sigma^2(t)=\sqrt{\frac{\mu}{a}}\), so long as
\(\bar x(0)\in\mathbb{R}\), \(\sigma^2(0)\in[0,+\infty)\),
\(N(0)\in(0,+\infty)\) and \(r>\frac{1}{2}\sqrt{a\mu}>0\). We prove
these statements in SM \S\ref{equilib}. Note in particular the
coincidence of equilibrial phenotypic variance predicted by this model
and that predicted by the classic Gaussian allelic model of quantitative
genetics (Bürger, 2000; Johnson and Barton, 2005; Lande, 1975). In
\S\ref{coev} and SM \S\ref{diffuse} we find this fitness function is
closely related to a fitness function derived from niche theory that we
use to model coevolution driven by resource competition.

@@

\begin{equation}
\|f\|_t=\sqrt{\mathbb{E}\left(\int_0^t\int_\mathbb{R}f^2(x,s)dxds\right)},
\end{equation} and make use of the convention \(f=g\) if \(\|f-g\|_t=0\)
for all \(t\geq0\). Later, when we investigate stochastic dynamics of
the abundance density \(\nu(x,t)\), we will find
\(\sqrt\nu\in\mathscr{N}_2\). This enables us to utilize the heuristics
developed in this section for the derivation of SDE describing the
stochastic dynamics of \(N(t)\), \(\bar x(t)\) and \(\sigma^2(t)\).

@@

\begin{equation}\label{exp_WN}
\mathbb{E}\big(\mathbf W_t(f)\big)=\mathbb{E}\big(\mathbf W_t(g)\big)=0,
\end{equation} \begin{equation}\label{cov_WN}
\mathbb{C}\big(\mathbf W_{t_1}(f),\mathbf W_{t_2}(g)\big)=\mathbb{E}\left(\int_0^{t_1\wedge t_2}\int_\mathbb{R} f(x,s)g(x,s)dxds\right),
\end{equation} where \(t_1\wedge t_2=\min(t_1,t_2)\) and \(\mathbb{C}\)
denotes covariance with respect to the underlying probability space. In
particular, denoting \(\mathbb{V}\) the variance operator with respect
to the underlying probability space, we have
\(\mathbb{V}\big(\mathbf W_t(f)\big)=\|f\|_t^2\) for all \(t\geq0\) and
\(f\in\mathscr{N}_2\). The operators \(\mathbb{E}\) and \(\mathbb{C}\)
are to be distinguished from \(\bar f(t)\) and \(\mathrm{Cov}_t(f,g)\)
which denote expectation and covariance with respect to phenotypic
diversity at time \(t\geq0\).

@@

\begin{equation}
\|I_D\|_t^2=\mathbb{E}\left(\int_0^t\int_\mathbb{R}I_D^2(x,s)dxds\right)=t\int_Ddx=t|D|<+\infty, 
\end{equation} where \(|D|\) denotes the length of \(D\), we have
\(I_D\in\mathscr{N}_2\). Thus, using equations (\ref{exp_WN}) and
(\ref{cov_WN}) and adopting the informal notation introduced in equation
(\ref{informal}), we can write the following

@@

\begin{equation}
\mathbb{E}\left(\int_0^td\hat{\mathbf W}_s(f)\right)=0, \ \mathbb{C}\left(\int_0^{t_1}d\hat{\mathbf W}_s(f),\int_0^{t_2}d\hat{\mathbf W}_s(f)\right)=t_1\wedge t_2
\end{equation} and in particular, as a function of \(t\),
\(\int_0^td\hat{\mathbf W}_s(f)\) is a standard Brownian motion for any
\(f\in\mathscr{N}_2\). Hence, \(d\hat{\mathbf W}_t(f)\) is analogous to
the traditional shorthand used to denote stochastic differentials. Thus,
equation (\ref{cov_WN_xi}) effectively extends Itô's multiplication
table to:

@@

\paragraph{Branching Brownian motion (BBM)}

A BBM tracks individuals navigating \(d\)-dimensional Euclidean space
that reproduce and senesce between exponentially distributed intervals.
Unlike other stochastic processes that take values in \(\mathbb{R}^d\),
BBM takes values in the set of \emph{non-negative finite measures} over
\(\mathbb{R}^d\). Intuitively, one can think of a finite measure as a
function that maps subsets of \(\mathbb{R}^d\) to real numbers. To be
formal, we only consider the Borel subsets of \(\mathbb{R}^d\)
corresponding to the Euclidean metric, but understanding this
technicality is not crucial to our discussion. In particular, denoting
\(X_t\) a BBM, for a subset \(D\subset\mathbb{R}^d\), \(X_t(D)\) returns
the (random) number of individuals alive within the region \(D\) at time
\(t\geq0\). The BBM has three main components:

@@

\paragraph{Interacting superprocesses}

In the above subsection we reviewed the diffusion limit of an especially
tractable measure-valued branching process. However, we found the
simplicity of this process restricts us from modelling nonlocal effects
on the fitness of individuals. Here, we discuss superprocesses that
account for such effects. The existence of diffusion limits for a class
of measure-valued branching processes involving nonlocal effects has
been treated by Méléard and Roelly (1993, 1992). The interactions can
manifest as dependencies of the spatial movement or reproductive law of
individuals on their position \(x\) and the state of the whole
population described by \(X_t\). An important result of Méléard and
Roelly (1993, 1992) is a theorem that provides sufficient conditions to
construct a sequence of rescaled measure-valued branching processes that
converge to a generalization of the Dawson-Watanabe superprocess that
includes interactions. The rescaling is analogous to that described
above for non-interacting Dawson-Watanabe superprocesses, but now the
reproductive law described by \(\mathscr{W}(X_t,x)\) and \(V(X_t,x)\),
branching rate \(\lambda(X_t,x)\) and diffusion parameter
\(\sqrt{\mu(X_t,x)}\) are allowed to depend on the whole population
\(X_t\) and individual location \(x\). In Figure \ref{rescaled} we
demonstrate this rescaling in discrete time for a population
experiencing stabilizing selection and logistic growth. Since time is
discretized, the process we simulate is formally a branching random
walk. For further details on our simulation see SM \S\ref{numerical}.

@@

\begin{multline}
N(t)-N(0)=\\
\int_0^t\int_\mathbb{R}\nu(x,s)\left(m(\nu,x)\cdot1+\frac{\mu}{2}\frac{\partial^2}{\partial x^2}1\right)+1\sqrt{V\nu(x,s)}\dot W(x,s)dsdx \\
=\int_0^t\bar m(s)N(s)dt+\int_0^t\sqrt{VN(s)}d\hat{\mathbf{W}}_s(\sqrt{\nu(x,s)}),
\end{multline} where

\begin{equation}
\bar m(t)=\frac{1}{N(t)}\int_\mathbb{R}m(\nu,x)\nu(x,t)dx,
\end{equation} and

@@

\begin{subequations}\label{gen_eqns}
\begin{equation}\label{N}
dN(t)=\bar m(t)N(t)dt+\sqrt{V N(t)}dW_1(t),
\end{equation}
\begin{equation}\label{xbar_gen}
d\bar x(t)=\mathrm{Cov}_t\Big(x,m(\nu,x)\Big)dt+\sqrt{V\frac{\sigma^2(t)}{N(t)}}dW_2(t),
\end{equation}
\begin{multline}\label{sig2_gen}
d\sigma^2(t)=\mathrm{Cov}_t\Big((x-\bar x(t))^2,m(\nu,x)\Big)dt+\left(\mu-V\frac{\sigma^2(t)}{N(t)}\right)dt\\+\sqrt{V\frac{\overline{(x-\bar x(t))^4}-\sigma^4(t)}{N(t)}}dW_3(t),
\end{multline}
\end{subequations}

@@

\begin{equation}\label{covxm}
\mathrm{Cov}(x,m)=\sigma^2\left(\frac{\partial\bar m}{\partial\bar x}-\overline{\frac{\partial m}{\partial\bar x}}\right),
\end{equation} \begin{equation}
\mathrm{Cov}\Big((x-\bar x)^2,m\Big)=2\sigma^4\left(\frac{\partial\bar m}{\partial\sigma^2}-\overline{\frac{\partial m}{\partial\sigma^2}}\right)
\end{equation} and \(\overline{(x-\bar x)^4}=3\sigma^4\). In particular,
these results imply

@@

\paragraph{Description of abiotic selection and competition}

@@

\begin{multline}
\mathcal O_{ij}(x_i-x_j)=\int_{-\infty}^{+\infty}u_i(\zeta,x_i)u_j(\zeta,x_j)d\zeta\\=\frac{U_iU_j}{\sqrt{2\pi(w_i+w_j)}}\exp\left(-\frac{(x_i-x_j)^2}{2(w_i+w_j)}\right).
\end{multline}

\paragraph{Avoiding caveats of niche theory}

@@

\paragraph{The fitness function}

To map the degree of niche overlap to fitness, we assume competition
between individuals with niche locations \(x_i\) and \(x_j\) decreases
the expected reproductive output for the individual in species \(i\) at
the rate \(c_i\mathcal{O}_{ij}(x_i-x_j)\) for some \(c_i>0\). We refer
to \(c_i\) as the sensitivity to competition for species \(i\). The term
\(c_i\mathcal{O}_{ij}(x_i-x_j)\) coincides with a special case of a term
used to capture competition in Dawson's geostochastic logistic model, an
SPDE model developed to study the combined effects of demographic
stochasticity, spatial dispersion and locally finite carrying capacity
(Dawson, 1978). In relation to the example fitness function discussed in
\S\ref{deterministic}, consider
\(\kappa(x_i-x_j)=\mathcal{O}_{ij}(x_i-x_j)\).

\paragraph{The model}

In SM \S\ref{diffuse} we combine our treatment of resource competition
with equations (\ref{N}), (\ref{xbarfinal}) and (\ref{Gfinal}) to find

@@

\paragraph{Description of model behavior}

@@

\paragraph{Natural history of an in-silico community}

@@

\begin{multline}
\alpha_{ij}(t)=\frac{c_i}{r_i(t)}\int_{-\infty}^{+\infty}\int_{-\infty}^{+\infty}p_i(x,t)p_j(y,t)\mathcal O_{ij}(x,y) dxdy \\ =\frac{c_iU_iU_j}{r_i(t)}\sqrt\frac{b_{ij}(t)}{2\pi}\exp\left(-\frac{b_{ij}(t)}{2}\big(\bar x_i(t)-\bar x_j(t)\big)^2\right),
\end{multline} where

@@

\paragraph{Metric of pairwise coevolution}

Below we investigate the correspondence of interaction intensity and
coevolutionary change. However, we can already identify one major
discrepancy; \(\alpha_{ij}\) is maximized when \(\bar x_i=\bar x_j\),
but \(\beta_{ij}=0\) under the same condition. We therefore include in
our metric of selection the standardized stabilizing selection gradient
\(\gamma\) which measures the effect of stabilizing or disruptive
selection on phenotypic variance (Lande and Arnold, 1983). In SM
\S\ref{diffuse} we show that the standardized stabilizing selection
gradient induced by species \(j\) on species \(i\) can be computed as

@@

\paragraph{Covariance of selection and competition as a function of diversity}

We now make use of the expressions derived for competition coeffecients
and selection gradients to investigate their relationship. As a first
pass, let us assume all model parameters are equivalent across species
and that each species has the same abundance and trait variance. Let us
further assume that richness \(S\) is large and the distribution of mean
trait values is normal with mean \(\bar{\bar x}\), variance
\(V_{\bar X}\) and density \(f_{\bar X}\). Such assumptions are typical
when deriving analytical results in the field of theoretical
coevolutionary community ecology (Nuismer et al., 2012, 2018). If
\(\bar{\bar x}\) is near \(\theta\) and \(V_{\bar X}\) is much smaller
than \(|2\frac{R}{a}-G-\eta|\), then we may approximate \(r_i\) with

@@

{\centering \includegraphics[width=1\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/cov}

@@

{\centering \includegraphics[width=0.6\linewidth]{/home/bb/Gits/branching.brownian.motion.and.spde/Current Version/Cαℭp}

@@

To understand patterns found in ecological communities considerations
must push beyond microevolutionary and contemporary ecological processes
and consider the macroevolutionary dynamics of ancestral lineages
leading to extant populations. Using sub-alpine flower communities as an
example, one can observe a very strict ordering of phenology across a
broad geographic range. In particular, whether in the Colorado Rocky
mountains (such as Gothic, Colorado) or on an outlier of the Idaho
batholith (such as Kamiak butte near Palouse, Washington), one will
almost surely observe a very conspicuous order of flowers emerging in
early spring: at the very beginning of the season blooms \emph{Claytonia
lanceolata} followed by \emph{Erythronium grandiflorum} which precedes
\emph{Delphinium nuttallianum} (B. Week, personal observations). If
contemporary phenological coevolution is rampant, why should this
pattern be so well preserved across a thousand miles of rugged and
diverse terrain? Although it would be exciting to find that these
species repeatedly coevolved this pattern in each location, a more
parsimonious hypothesis suggests the phenology and physiology of these
species slowly evolved independently over macroevolutionary time scales
to take advantage of the specific conditions available within each of
these windows of the flowering season. However, considering genera of
these species have many early-season representatives occurring further
east, such as in the Appalachian Mountains, these adaptations
potentially arose before the glaciers of the Pleistocene began to
retreat (Paul CaraDonna \& Dave Tank, personal communications). Hence,
given these considerations, it appears that an understanding of early
season phenology patterns of sub-alpine flower communities should focus
on how these communities are assembled as opposed to contemporary
evolutionary dynamics. Indeed, recent work testing phylogeographic
models suggests alpine flower communities tend to follow neutral
assembly where flowers merely compete for who can disperse to new
habitat first (Marx et al., 2017). Due to spatial and altitudinal
proximity, a similar result is likely to hold for sub-alpine flower
communities.
